An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-25 paper/5.experiments.tex
Create 5.experiments.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos

We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, Ocaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.

@@

We have conducted 11 test cases shown in Table \ref{tab:test-cases}. $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 11 functions mentioned in the table. Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

\begin{table}[hbt!]
    \centering
    \caption{Test cases}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|l @{\extracolsep{\fill}}|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Type}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{4-7} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
        1 & lcomp-replace & straightline & 1 & 100 & 100 & 1 \\ \hline
        2 & lcomp-replace-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
        3 & lcomp-phead-ldata & straightline & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        4 & lcomp-ptail & straightline & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
        5 & lcomp-ptail-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
        6 & lcomp-bsnoc & straightline & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
        7 & lcomp-bsnoc-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
        8 & bmapreplace & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] \\ \hline
        9 & bmapreplace-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
        10 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
        11 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
    \end{tabular*}
\end{table}

The first 7 cases maunally use $n$ compose operators to make straightline compositions from $n + 1$ similar pure programs. We use the term \textit{pure} here to refer to programs that are not self-compose. The prefix \textit{lcomp} and \textit{rcomp} denote the associate precedences of a manual composition. The suffix \textit{ldata} indicates that the input size is large. The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = \underbrace{[T,\ldots,T]}_{\text{n times}}$,\ $T = \underbrace{[A,\ldots,A]}_{\text{10 times}}$,\ $A = [1,\ldots,5]$

If the associative type and input size are omitted, the table, in general, will comprise 4 programs: composition of many \textit{phead}s, composition of many \textit{ptail}s, composition of many \textit{bsnoc}s and \textit{breverse}.

We introduced the composition $phead \circ phead$ earlier. The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view. We now review the behavior of the remaining 3 programs.

@@

\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function for \textit{foldr} which is an important higher-order function on lists. This case is more natural than the other since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}. It will look like a composition of more BX programs at some point if we slow down the evaluation. Note that $\circ$ has a higher priority than $\times$. This leads to seemingly impossible to transform \textit{bReverse} as well as \textit{bFoldr} from the left associative style to the right one.

\subsection{Results and discussions}

@@

Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 12 test cases. $put$ works more efficiently in the cases of right associative compositions, rather than the left ones. This makes sense because the number of \textit{get} when evaluating a right associative composition is linear in the number of compositions, while the respective number when computing a left associative composition is significantly larger. For self-composed program, \textit{breverse}, it could be impossible to change the associate precedence from left to right. Then, $put$ works less efficiently than both $put_m$ and $xpg$. In general, for left associative compositions, if the input size is not too large, $put_m$ is the fastest method. However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table. At that time, $xpg$ proved to be the most effective method.

>>>>

We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.

@@

We have conducted 7 test cases shown in Table \ref{tab:test-cases}. Although the number of programs is not so much, they are not trivial and can be completely used in practice. In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table. These $bx$ are left associative composition since we we are only interested in this kind of program in the paper. Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

\begin{table}[hbt!]
    \centering
    \caption{Test cases}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|l @{\extracolsep{\fill}}|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Type}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{4-7} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
        1 & lcomp-phead-ldata & straight line & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        2 & lcomp-ptail & straight line & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
        3 & lcomp-ptail-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
        4 & lcomp-bsnoc & straight line & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
        5 & lcomp-bsnoc-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
        6 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
        7 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
    \end{tabular*}
\end{table}

The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from $n + 1$ similar \textit{pure} programs. We use the term \textit{pure} here to refer non-recursion compositions. The prefix \textit{lcomp} means that the associate precedence of $\circ$ in a straight-line composition is left. The suffix \textit{ldata} indicates that the input size is large. The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$. These structures are only intended to generate data enough large for observing experimental results.

We introduced the composition $phead \circ phead$ earlier. The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view. Because of the type of the source, this program is seen as a \textit{ldata} case. Next, we briefly explain the behavior of the remaining compositions in table \ref{tab:test-cases}.

@@

\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function of \textit{foldr} which is an important higher-order function on lists. This program, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list. \textit{breverse} is more natural than the others since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}. It will look like a composition of more programs at some points in the evaluation if we slow down this process. Note that $\circ$ has a higher priority than $\times$. This leads to seemingly impossible to transform \textit{bReverse} as well as \textit{bFoldr} from the left associative style to the right one.

\subsection{Results}

Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases. We simply ignore the experimental results of $pg$, $cpg$ and $kpg$ since they are shown the slowness compared to $xpg$. For left associative compositions, $put$ works poorly because of the number of reevaluated $get$s. The figure shows $put_m$ and $xpg$ are better than $put$ in some cases. The left of figure contains tests using not very big input, and the right includes of the others. We see that $put_m$ is the fastest method if the input size is not too large. However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table. At that time, $xpg$ proved to be the most effective method.

@@
